## Flexible GLMs

Now that we can write `brms` models and read their output, we can brave more
complicated models.

```{r Load libraries}
#| code-summary: "Load libraries"
#| code-fold: true
#| eval: true
#| output: false
library(ggplot2)
library(brms)
library(tidybayes)
library(bayesplot)
options(brms.backend = "cmdstanr")
```

Let us consider some data from campers gathered by UCLA[^orig-pois]. In their words "the
state wildlife biologists want to model how many fish are being caught by
fishermen at a state park. Visitors are asked whether or not they have a
camper, how many people were in the group, were there children in the group and
how many fish were caught. Some visitors do not fish, but there is no data on
whether a person fished or not. Some visitors who did fish did not catch any
fish [...]"[^ucla-data]

[^orig-pois]: This data analysis is adapted from Paul Bürkner's original, available in <https://cran.r-project.org/web/packages/brms/vignettes/brms_distreg.html>.

[^ucla-data]: This description comes from <https://stats.oarc.ucla.edu/stata/dae/zero-inflated-poisson-regression/>, though it is unclear if the data are real
or simulated

These data describe 250 groups of park visitors. For each group, we know how
many fish they caught (`count`), how many total people were in the group
(`persons`), how many of these were children (`child`), and whether the group
brought a camper (`camper`).

```{r Prepare fisher data}
#| eval: true
fisher_df <- read.csv("https://paul-buerkner.github.io/data/fish.csv")
fisher_df <- fisher_df[, c("persons", "child", "camper", "count")]
names(fisher_df)[names(fisher_df) == "count"] <- "num_fish"
summary(fisher_df)
```

It is reasonable to think that larger groups were more likely to catch more
fish. Also, bringing a camper signals that people are staying longer at the
park, which would give them more time to fish. But children do not often like
to sit silently for long periods, so groups with more children may have fished
less.

These ideas suggest using a model that can associate the average number of
caught fish to the `persons`, `child`, and `camper`. We should also use a model
that accounts for the fact that the number of fishes caught is a non-negative
integer. A Poisson regression is a relatively simple model that can fulfil both of these requirements.

Note, however, that many visitor groups caught zero fishes (see figure below). These zeroes are
a combination of those visitors that did not fish at all and those that did
but caught nothing. Unfortunately, there are many more zeroes than any regular
Poisson regression can account for. Fortunately, with `brms` we can expand our
model to be able estimate the proportion of groups that did not fish at all.
This expanded model is called "zero-inflated Poisson regression" because it
inflates (i.e., adds more) zeroes to the basic Poisson regression. For now, we presume that all groups had the same chance of going fishing.

```{r histogram of fishes caught}
#| code-fold: true
#| eval: true
#| warning: false
#| fig-align: center
#| fig-width: 3.5
#| fig-height: 3
ggplot(data = fisher_df, mapping = aes(x = num_fish)) +
  geom_histogram(binwidth = 2, color = "black", fill = "navyblue") +
  labs(
    title = "Many groups did not catch any fish",
    x = "Number of fish caught",
    y = "Number of groups"
  ) +
  theme_bw()
```

The code to fit this zero-inflated regression is shown below. The only novelty
here is the argument `family`, which controls the type of model we want to fit---in this case, a `zero_inflated_poisson()`.

```{r fit zero inf poisson}
fit_fisher1 <- brm(
  num_fish ~ persons + child + camper,
  data = fisher_df,
  family = zero_inflated_poisson(),
  chains = 4,
  cores = 4,
  iter = 1000,
  warmup = 500
)
```



```{r summary fit_fisher1}
summary(fit_fisher1)
```

$$
\begin{align*}
    y&: \text{Number of fish}\\
    y&\sim \text{Poisson}(\mu) \\
    \text{log}(\mu) &= \beta X \\
    \beta_j &\sim \mathrm{i.i.d.} \text{ Normal}(0, 1) \text{ for all } j
\end{align*}
$$