## Flexible GLMs

Now that we can write `brms` models and read their output, we can brave more
complicated models.

```{r Load libraries}
#| code-summary: "Load libraries"
#| code-fold: true
#| eval: true
#| output: false
library(ggplot2)
library(brms)
library(tidybayes)
library(bayesplot)
options(brms.backend = "cmdstanr")
```

Let us consider some data from campers gathered by UCLA. In their words "the
state wildlife biologists want to model how many fish are being caught by
fishermen at a state park. Visitors are asked whether or not they have a
camper, how many people were in the group, were there children in the group and
how many fish were caught. Some visitors do not fish, but there is no data on
whether a person fished or not. Some visitors who did fish did not catch any
fish [...]"[^ucla-data]

[^ucla-data]: This description comes from https://stats.oarc.ucla.edu/stata/dae/zero-inflated-poisson-regression/, though it is unclear if the data are real
or simulated

These data describe 250 groups of park visitors. For each group, we know how
many fish they caught (`count`), how many total people were in the group (`persons`), how many of these were children (`child`), and whether the group
brough a camper (`camper`).

```{r Prepare data}
fisher_df <- read.csv("https://paul-buerkner.github.io/data/fish.csv")
fisher_df <- fisher_df[, c("persons", "child", "camper", "count")]
summary(fisher_df)
```

