# Getting started with brms

## Installation

Making `brms` work in our computer requires some preparation. First, updating
R to its latest version will help our libraries work better with each other
and with R itself. Installing brms with an older version of R may work, but is
likely to cause problems later.

### C++ toolchain

`brms` uses C++ to translate our models in R into a machinespeak that the
computer runs more efficiently. So, our computer needs a toolchain to run C++
in the background. Setting up this toolchain can take several minutes, but
needs to be done only once.

**In Windows**, the C++ toolchain needs *RTools* 4.5. To check if this is
already installed, go to R and run `devtools::find_rtools()` (you may need to
first run `install.packages("devtools")`). To install RTools, first go to
[this website](https://cran.r-project.org/bin/windows/Rtools/rtools45/rtools.html).
Here, find the *Rtools45 installer*, download it, and follow its instructions
while keeping all the default choices.

**In macOS**, the C++ toolchain needs *Xcode Command Line Tools*. To check if
this is already installed, go to the Terminal (open `Finder > Applications >
Terminal`) and run
```
g++ --version
make --version
```

If your mac has `g++` at version 4.9.3 or later and `make` at version
3.81 or later, you do not need to install anything else (although newer
versions of `g++` are faster). If you do not have these tools, or if you want
to update them, run

```
sudo apt install g++
sudo apt install make
```

### Default version of brms

With our C++ toolchain ready, we can go back to R and install `brms`. It may take a few minutes to download and install all the packages bundled with `brms`.

```{r install brms}
#| eval: false
install.packages("brms")
```

We can verify that our `brms` installation works properly by running the
commands below. For now, you do not need to understand its meaning.

```{r test brms installation}
#| eval: false
library(brms)
fit1 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson())
summary(fit1)
```

The final part of your output should resemble this (the numbers can change
slightly):

```
Regression Coefficients:
           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept      1.77      0.12     1.53     2.00 1.00      642     1149
zAge           0.10      0.09    -0.07     0.27 1.00      690      873
zBase          0.70      0.12     0.46     0.95 1.00      591     1080
Trt1          -0.26      0.16    -0.58     0.05 1.00      699     1319
zBase:Trt1     0.05      0.17    -0.29     0.38 1.00      681     1115
```

::: {.callout-warning}
### Install toolchain before brms
R would not complain if we installed `brms` *before* preparing the C++
toolchain. But without the toolchain none of our models would run, so it is
better to set it up in advance
:::

### Faster brms with cmdstanr

`brms` uses a backend to interface with Stan. The default backend is `rstan`,
which does not require any additional installation but is somewhat slow. An
alternative backend is `cmdstanr`, which needs to be installed separately but is much faster than `rstan`. The extra setup is well worth the effort.

To install `cmdstanr`, we first download the package in R:

```{r}
#| eval: false
# Run this in a fresh R session or after restarting your current session
install.packages(
    pkgs = "cmdstanr",
    repos = c('https://stan-dev.r-universe.dev', getOption("repos"))
)
```

Now we can verify that the C++ toolchain also works with cmdstanr

```{r check cmdstan toolchain}
#| eval: false
library(cmdstanr)
check_cmdstan_toolchain()
```
```
The C++ toolchain required for CmdStan is setup properly!
```

Now we can ask R to configure `cmdstanr` for us. This process can take a few
minutes and will show many weird messages. We can ignore these messages as
long as R does not show an error.

```{r}
#| eval: false
install_cmdstan(cores = 2)
```

We can verify that `brms` can now call `cmdstanr` by running

```{r test brms with cmdstanr}
#| eval: false
library(brms)
options(brms.backend = 'cmdstanr') # This forces brms to use cmdstanr
fit1 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson())
summary(fit1)
```