# Linear regression in brms

`brms` allows us to add or remove parts of a model as needed, so our code is
only as complicated as the model we want to fit. Simple code for simple
models; fancy code for fancy models.

Imagine, for example, that we like penguins and we want to study their weight.
R has some clean data about three types of chilly fellows: Adelie, Chinstrap,
and Gentoo.

```{r load penguin data}
#| eval: true
library(ggplot2)
data(penguins)
penguins$body_mass_kg = penguins$body_mass / 1000
```

To start, let us compare the weights of these three groups. The histogram
below suggests that Gentoo penguins tend to be the heaviest, while
Adelie and Chinstrap tend to have similar weights.

```{r visualize penguin weights}
#| code-fold: true
#| eval: true
#| warning: false
#| fig-align: center
#| fig-width: 4
#| fig-height: 4
ggplot(penguins, mapping = aes(x = body_mass_kg)) +
    geom_histogram(binwidth = 0.2, color = "black", fill = "darkorange") +
    facet_wrap(~species, nrow = 3, ncol = 1) +
    labs(
    title = "Gentoo penguins are the heaviest",
    x = "Body mass (kg)",
    y = "Count"
    ) +
    theme_minimal()
```

We can complement this histogram with estimates for average weight of each
group of penguins. A linear regression can give us these estimates along with
their confidence intervals. We can fit this (and any other) regression in
`brms` using the function `brm()` (short for "bayesian regression model"),
which for our simple model requires only a formula to describe the regression
function and the name of the data set that houses the corresponding variables.

The code below shows how to fit a linear regression with weight (in kg) as the
dependent variable and species as the independent categorical variable with
values "Adelie", "Chinstrap", and "Gentoo". Run this code and ignore its
lengthy output for now.

```{r simple model with brm}
library(brms)
options(brms.backend = "cmdstanr")
fit_peng1 <- brm(
    formula = body_mass_kg ~ species,
    data = penguins
)
```

The first time it sees a model, `brm()` compiles the code, i.e., it uses the
C++ toolchain to translate the code into efficient machinespeak. So, this first
run seemed slow a few seconds, but rerunning this same model will be much
faster (try it!). Think of `brm()` as a virtual chef that can cook almost
anything you want but has to prepare the kitchen *after* you tell it what you
want.

By default, `brms` prints a lot of information related to the background work
it does to fit models, including progress reports that signal how much more
time it takes to run. In the code above, `refresh = 0` omits these reports.