# Linear regression in brms

It is good to approach `brms` as a modular library that helps us add or swap
parts of a model as needed. With this modularity, we need not resort to
subtle syntax or elaborate structures; we use simple code for simple models.

Imagine, for example, that we want to know the average weight of three types of penguins: Adelie, Chinstrap, and Gentoo. R has some data about these penguins.

```{r prepare penguin data}
#| eval: true
# Load and prepare data ####
data(penguins)
penguins$body_mass_kg = penguins$body_mass / 1000
summary(penguins)
```

```{r visualize penguin weights}
#| eval: true
library(ggplot2)
ggplot(penguins, mapping = aes(x = body_mass_kg)) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~species, nrow = 3, ncol = 1) +
  labs(title = "Gentoo penguins are the heaviest") +
  theme_bw()
```

Let us fit a regression to estimate the average weight of our three species.
Base R includes the widely-used `lm()` function. For this simple model, `lm()` requires a formula to describe the regression function, and the name of the
data set to find the corresponding variables.

```{r simple model with lm}
fit_peng_boring <- lm(
    body_mass_kg ~ species,
    data = penguins
)
summary(fit_peng_boring)
```

The code to fit the same model using `brms` is equally simple. Do not try to
run this for now unless you have previously installed `brms`.

```{r simple model with brm}
library(brms)
options(brms.backend = "cmdstanr")
fit_peng <- brm(
    body_mass_kg ~ species,
    data = penguins,
    refresh = 0 # To omit progress reports
)
```